---
title: "Incarceration Trends Data Visualization Analysis"
author: "Vincent Li"
date: '2022-11-23'
output: html_document
---

```{r echo=FALSE, eval=TRUE, message=FALSE, warning = FALSE}
source("../source/a4-helpers.R")
source("../source/analysis.R")
```

## Abstract 
  This report analyze ... summarize.. visualize... 

## Introduction 
  The dataset is provided by the project of _Vera_ that includes data collections of the Census of Jails (COJ) and the Annual Survey of Jails (ASJ) from 2013 to 2020. 

## Data Summary 
``` {r echo = TRUE}
highest_proportion_black_2013

```
This value

``` {r echo = TRUE}
highest_change_proportion_black

```
This value

``` {r echo = TRUE}
highest_proportion_black_2018

```
This value

``` {r echo = TRUE}
highest_change_proportion_black

```
This value

## Growth of the U.S. Prison Population
``` {r echo = TRUE}
get_year_jail_pop()
plot_jail_pop_for_us()

```
This chart

## Growth of the Prison Population by State
``` {r echo = TRUE}
check_states <- c("WA", "CA", "NY", "OR", "KS")

get_jail_pop_by_states(check_states)

plot_jail_pop_by_states(check_states)

```
This chart

## Change in Black People Jail Proportion in Mobile County (1970-2018)
``` {r echo = TRUE}
get_jail_prop_year_Mobile()

plot_jail_prop_year_Mobile()

```
This chart 

## Proportion of Black People in Jail Populatoin Each State 2013
``` {r echo = TRUE}
get_state_black_jail_prop()

map_black_jail_prop_states()

```
This map 

